---
title: "classification_practice"
output: html_notebook
---

```{r}
# Load packages
library(ggplot2)
library(tidyverse)
library(corrplot)
library(fastDummies)
```

```{r}
# Load the dataset
dataset <- read_csv("bank-data.csv")

# Inspect the first 15 entries
head(dataset, 15)
```

```{r}
# check the proportion of subsribed and not subsribed customers
table(dataset$subscribed) %>% prop.table()
```

```{r}
# find out if there is any missing values
any(is.na(dataset))

# find out how many rows with missing values
sum(is.na(dataset))
```

```{r}
# find out the columns having missing values
names(which(sapply(dataset, FUN=anyNA)))
```

```{r}
table(dataset$age)
```

```{r}
# Plot a bar chart of the counts of the customer age
barplot(table(dataset$age))
```

```{r}
boxplot(dataset$age)
```

```{r}
hist(dataset$age, xlim = c(0,100), breaks=100)
```

```{r}
boxplot(age~subscribed, data=dataset)
```

```{r}
younger <- subset(dataset, age < 60)

boxplot(age~subscribed, data = younger)
```

```{r}
boxplot(numberTimesContactedInCampaign~subscribed, data = subset(dataset, numberTimesContactedInCampaign <= 10))
```

```{r}
boxplot(daysSinceLastContact~subscribed, data = dataset)
```

```{r}
boxplot(totalTimesContacted~subscribed, data = dataset)
```


```{r}
subset_1 <- subset(dataset, subscribed == "yes")
subset_2 <- subset(dataset, subscribed == "no")

hist(subset_1$day_of_month)
hist(subset_2$day_of_month)
```

```{r}
hist(dataset$numberTimesContactedInCampaign)
```

```{r}
levels(factor(dataset$subscribed))
```


```{r}
# Convert any variables you want to test as numeric, using "as.numeric()" function, and save these as new vectors (e.g., dataset_subscribed, dataset_age, etc.)
dataset_subscribed <- as.numeric(as.factor(dataset$subscribed))
dataset_age <- dataset$age
dataset_job <- as.numeric(as.factor(dataset$job))
dataset_dom <- as.numeric(as.factor(dataset$day_of_month))
dataset_ntcin <- dataset$numberTimesContactedInCampaign
dataset_last_contact <- dataset$daysSinceLastContact
dataset_ttc <- dataset$totalTimesContacted
# Once your data has been changed from categorical to numeric, you can now perform a correlation matrix. Save the newly converted variables into a new data frame called "corr_data" (this makes future steps easier by excluding any data you do not need);
corr_data<-data.frame(dataset_subscribed, dataset_age, dataset_job, dataset_dom, dataset_ntcin, dataset_last_contact, dataset_ttc)
# rename the variables in your new data frame by using the setNames() function:
corr_data<-setNames(corr_data, c("subscribed", "age", "job", "day_of_month", "numberTimesContactedInCampaign", "daysSinceLastContact", "totalTimesContacted"))
# check headers to make sure the variables in the new data frame are named correctly
head(corr_data)
# run the correlation matrix algorith on your "corr_data" dataframe, by using the cor() function, and rename it as a new vector called "corr_matrix"
corr_matrix <- cor(corr_data)
corr_matrix
# install and load the R package "corrplot" so you can graph your new correlation matrix vector:

corrplot(corr_matrix, method="pie", type='upper')
```

```{r}
fastDummies::dummy_cols(dataset, select_columns = "job")
```

### Understanding *p*-values.
The p-value is the probability of finding the same result for something if you repeated the same test/analysis 100 times,

If this probability is lower than the conventional 5% (*p*<0.05) the correlation coefficient is called statistically significant, meaning the probability of this event or finding occurring again is not due to chance (i.e., a less than %5 chance this occurrence is random/due to chance).

In relation to correlation coefficients, a *p*-value of <0.05 tells us that there is a <5% chance that there is no correlation between two parameters (i.e., the probability of a true correlation is high). Calculate some correlations between parameters. A *p*-value <0.05 is a good correlation

```{r}
# this is how to statistically test (p-value) the correlation between two variables
# testing correlation between job and subscribed
cor.test(dataset_job, dataset_subscribed, method=c("pearson"))
# testing correlation between marital status and subscribed
#cor.test(as.numeric(dataset$marital), dataset_subscribed, method=c("pearson"))
# testing correlation between level of education and subscribed
#cor.test(as.numeric(dataset$education), as.numeric(dataset$subscribed), method=c("pearson"))
# testing correlation between housing and subscribed
#cor.test(as.numeric(dataset$housing), as.numeric(dataset$subscribed), method=c("pearson"))
```
What can we conclude about these correlations based on their p-values?
```

