---
title: "Clustering Exercise - HK District Data"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
# Load packages
library(tidyverse)
```

The dataset for this exercise is from [Hong Kong 2016 Census](https://www.bycensus2016.gov.hk/en/). It lists the average income and median age under different district councils in Hong Kong.

## 1. Prepare and Explore Data
```{r}
# 2. Load data file
df <- read_csv("district_data.csv")

# check first a few rows
head(df)
```

```{r}
# check the dimension of data
dim(df)
```


```{r}
# check the summary of data
summary(df)
```

## 2. Data Cleaning
```{r}
# check column names
names(df)

# rename the column names
names(df) <- c("district", "avg_income", "median_age")
names(df)
```

The data type of `avg_income` column is string now, so we need to convert it to numeric.
```{r}
# remove $ and thousand separator
df$avg_income <- gsub("[$,]", "", df$avg_income)

# convert to numeric
df$avg_income <- as.numeric(df$avg_income)

# check the column types
glimpse(df)
```

## 3. Data Visualization
```{r}
qplot(x = df$median_age, y = df$avg_income, label = df$district, geom = "text")
```

## 4. K-means Clustering
Because the income data is in thousand, much 'bigger' than age data, so we need to scale the income data before clustering.

```{r}
df_scaled <- as.data.frame(scale(df[c("median_age", "avg_income")]))

head(df_scaled)
```

```{r}
# define the number of clusters
k_clusters <- 3

# get clustering result
k_means_result <- kmeans(df_scaled, centers = k_clusters)

# add cluster result to original data
df$cluster <- as.factor(k_means_result$cluster)

head(df)
```

```{r}
# visualize cluster result
qplot(x = df$median_age, y = df$avg_income, label = df$district, geom = "text", color = df$cluster)
```

## 5. Decide the optimal number of clusters
```{r}
# set maximum number of cluster
k_cluster_max <- 10
```

```{r}
# iterate through the number of clusters from 1 to 10
within_sum_of_squares <- sapply(1:k_cluster_max, 
                                function(k){
                                  kmeans(df_scaled, centers = k)$tot.withinss})

# visualize the output
plot(1:k_cluster_max, within_sum_of_squares, type = "b",
     xlab = "Number of clusters", ylab = "Total within cluster sum of squares")
```

The chart shows that probably 4 is the correct number of clusters as the curve bents most at this point, so we repeat clustering with 4 clusters.

```{r}
# conduct clustering with 4 clusters
k_means_result_2 <- kmeans(df_scaled, centers = 4)

# add result to dataframe
df$cluster <- as.factor(k_means_result_2$cluster)
```

```{r}
# visualize cluster result with ggrepel to avoid label overlapping
library(ggrepel)
qplot(x = df$median_age, y = df$avg_income, label = df$district, color = df$cluster,
      xlab = "Median Age", ylab = "Average Income") + geom_text_repel()
```

The graph shows that the districts has been allocated to 4 groups, the result is reasonable and reliable.