---
title: "Week4_Challenge"
output: pdf_document
---

This notebook is to complete the challenge of `Week 4 Data`.

### Goal of the module
By the end of this module, you will have loaded, cleaned and analysed different datasets using R.

### Learning Objectives
- I understand different data structures
- I am able to load different types of files in R
- I know the difference between .csv, .xlxs, .tsv and .json files
- I can evaluate and solve the different problems of "dirty" data
- I can prepare visual and analytical data analysis
- I am able to filter datasets base on columns and row

### Create a section to load revelant packages
```{r}
library(tidyverse) # data manipulation
library(readxl) # read xls file
library(httr) # work with URL and HTTP
library(gdata) # data manipulation
library(jsonlite) # work with json file
library(stringr) # work with strings
library(GGally) # extension of ggplot2
```

## Loading data
```{r}
# use read.csv() to load data and store it to a variable
happiness <- read.csv("data/introduction_to_data/happiness.csv")

# check first 7 rows
head(happiness, 7)
```

```{r}
# use read_excel() to load data and store it to a variable
df_xlsx <- read_excel("data/introduction_to_data/happiness.xlsx")

# check the last 7 rows
tail(df_xlsx, 7)

# check dimension
dim(df_xlsx)
```

```{r}
# use read.table() to load data and store it to a variable
df_tsv <- read.table(file = "data/introduction_to_data/happiness.tsv", header = TRUE, sep = "\t")

# check the first a few rows
head(df_tsv)
```

```{r}
# use fromJSON() to load data and store it to a variable
df_json <- fromJSON("data/introduction_to_data/grades.json")
df_json <- df_json$data

# check the last 6 rows
tail(df_json)
```

```{r}
# load GapMinder data
df_gm <- read_excel("data/introduction_to_data/gap_minder.xlsx", sheet = 1, )

# check first 6 elements
head(df_gm)
```

## Cleaning Data
### Cleaning Data for `happiness` database
```{r}
# check first 6 elements
head(happiness)

# replace column names
names(happiness) <- c('Country', 'Region', 'Hemisphere', 'HappinessScore', 'HDI', 'GDP_PerCapita', 'Beer_PerCapita', 'Spirit_PerCapita', 'Wine_PerCapita')

# check first 6 elements again
head(happiness)
```

```{r}
# check the data type of all variables with a single line of code
str(happiness)

# Alternative method
sapply(happiness, class)
```

```{r}
# replace commas in columns "HappinessScore" and "GDP_PerCapita" and change their data types to numeric
happiness$HappinessScore <- as.numeric(gsub(",", ".", happiness$HappinessScore))
happiness$GDP_PerCapita <- as.numeric(gsub(",", ".", happiness$GDP_PerCapita))

# check the first 6 elements to confirm the result
head(happiness)
```

```{r}
# correct typos and multiple categories of "Hemisphere" so that it only contains `north`, `south` and `both`
unique(happiness$Hemisphere)

# replace the typos and other categories to required words
happiness$Hemisphere <- gsub("^[Ss]\\w*", "south", happiness$Hemisphere, ignore.case = TRUE)
happiness$Hemisphere <- gsub("^[Nn]\\w*", "north", happiness$Hemisphere, ignore.case = TRUE)

# check the unique values of Hemisphere column again
unique(happiness$Hemisphere)

```

```{r}
# check the first 6 rows of the cleaned data
head(happiness)

```

### Cleaning Data for `gapminder_data` databse
```{r}
# check the data
head(df_gm)

```

```{r}
# Change the name of the variable 'Total 25-54 unemployment (%)' to "Country"
names(df_gm)[1] <- "Country"

```

```{r}
# "Unpivot" the database gapminder_data, so each observation will be described by three variables: Country, Year, Unemployment.
gmd_melt <- gather(df_gm, key = "Year", value = "Umemployment", -Country)

# check melted dataframe
head(gmd_melt)
```

```{r}
# Change the format of the variable Year to numeric
gmd_melt$Year <- as.numeric(gmd_melt$Year)

# check result
head(gmd_melt)
```

```{r}
# Drop any row that contains a NaN value using the function .dropna. Save that file into gdm_noNa
gdm_noNa <- drop_na(gmd_melt)

# check if it worked
head(gdm_noNa)
```

```{r}
# Check out how many elements you dropped using the function .shape
drop_row <- dim(gmd_melt)[1] - dim(gdm_noNa)[1]

print(drop_row)
```

## 4. Summarising Data
```{r}
# Count how many observations are in the happiness dataset using the function length(); save that number in a variable called totalCounts
totalCounts <- length(happiness)


# Calculate the sum of all the Happiness score for all countries, save that value in a variable called summation
summation <- sum(happiness$HappinessScore)


# Calculate the mean with the two numbers above
mean(totalCounts, summation)


# Calculate the arithmetic mean and the median of HappinessScore
hs_mean <- mean(happiness$HappinessScore)
print(paste("The arithmetic mean of HappinessScore is: ", hs_mean))

hs_median <- median(happiness$HappinessScore)
print(paste("The arithmetic median of HappinessScore is: ", hs_median))


# Calculate the variance, the standard deviation and the interquartile range of HappinessScore
hs_var <- var(happiness$HappinessScore)
print(paste("The variance of HappinessScore is: ", hs_var))

hs_std <- sd(happiness$HappinessScore)
print(paste("The standard deviation of HappinessScore is: ", hs_std))

hs_iqr <- IQR(happiness$HappinessScore)
print(paste("The interquartile of HappinessScore is: ", hs_iqr))


# Calculate the maximum and the minimum value of HappinessScore
hs_max <- max(happiness$HappinessScore)
print(paste("The maximum value of HappinessScore is: ", hs_max))

hs_min <- min(happiness$HappinessScore)
print(paste("The minimum value of HappinessScore is: ", hs_min))
cat("\n")


# In a single line of code, calculate the mean of all the numeric variables
print("The mean of numeric variables are as below:")
print(sapply(happiness[,4:9], mean, na.rm = TRUE))
cat("\n")


# In a single line of code calculate: (a) minimum value, (b) Q1, (c) median, (d) mean, (e) Q3, (f) maximum value.
summary(happiness)

```

```{r}
# Pivot a table to know the average value of all the numeric variables based on (a) Hemisphere and (b) Region. Save the results in `hemisfere` and `region`, respectively
# group data by hemisphere
gb_hemisphere <- group_by(happiness, Hemisphere)

# summarise the mean of numeric columns for all hemispheres
hemisphere <- summarise_if(gb_hemisphere, is.numeric, mean)
hemisphere
```

```{r}
# group data by region
gb_region <- group_by(happiness, Region)

# summarise the mean of numeric columns for all regions 
region <- summarise_if(gb_region, is.numeric, mean)
region
```


## 5. Data Visualisation
```{r}
# Create a vertical boxplot of all the HappinessScore variable from the dataset happiness, set the title of the the plot to Global Happiness Score, change the y axis to Happiness Score
boxplot(happiness$HappinessScore, main = "Global Happiness Score", ylab = "Happiness Score")

```

```{r}
# Create a vertical boxplot of all the alcohol consumption variables from the dataset happiness, set the title of the the plot to Country Alcohol Consumption, change the y axis to Litres per capita (l), and change the labels of the x axis to Beer, Spirits, Wine.
df <- gather(happiness, key = "Category", value = "Consumption", -c(Country, Region, Hemisphere))
df <- filter(df, df$Category == "Beer_PerCapita" | df$Category == "Spirit_PerCapita" | df$Category == "Wine_PerCapita")
```

```{r}
boxplot(Consumption~Category, df, main = "Country Alcohol Consumption", ylab = "Litres per capita (l)", xaxt = "n")
axis(1, at = 1:3, labels = c("Beer", "Spirits", "Wine"))

```
```{r}
# Create a horizontal boxplot of HappinessScore, separated by Hemispheres. Make the title Happiness Score in each hemisphere; change the labels in the y axis to north, both, and south. Change the labels in the x axis to Happiness Score
boxplot(HappinessScore~Hemisphere, happiness, horizontal = TRUE, main = "Happiness Score in Each Hemisphere", xlab = "Happiness Score", ylab = "Hemisphere")

```

```{r}
# Create a vertical violin plot of GDP_PerCapita, separated by Region, change the title and the labels accordingly
# draw the violin chart and fill in each region with color
p <- ggplot(happiness, aes(x = Region, y = GDP_PerCapita)) + geom_violin(scale = "width", aes(fill = Region)) + ggtitle("GDP Per Capita in Each Region")

# Rotate the x-tick labels and remove legend
p + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")


```

```{r}
# Create a scatter plot between HappinessScore (dependent variable) and Wine_PerCapita (independent variable)
ggplot(happiness, aes(x = Wine_PerCapita, y = HappinessScore, color = Hemisphere),
       main = "Wine Per Capita vs Happiness Score",
       xlab = " Wine Per Capita",
       ylab = "Happiness Score") + geom_point()
```

```{r}
# Check out the distribution of HappinessScore using a histogram
ggplot(happiness, aes(x = HappinessScore)) + geom_histogram(binwidth = 0.5, color = "white", fill = "lightblue") + ggtitle("Distribution of Happiness Score") # color for border color

```

```{r}
# Check out the distribution of HappinessScore using a density curve
# add basic density plot and change color
p2 <- ggplot(happiness, aes(x = HappinessScore)) + geom_density(fill = "lightblue", alpha = 0.4) + ggtitle("Distribution of Happiness Score")

# add mean line
p2 + geom_vline(aes(xintercept = mean(HappinessScore)), color = "darkblue", linetype = "dashed")
```

```{r}
# Create a plot which contains multiple scatter plots between all the variables and also contains the distribution of the numerical data, coloured by hemisphere
# use ggscatmat from ggally package
ggscatmat(happiness, color = "Hemisphere", alpha = 0.8)
```

## 6. Filtering Data
`happiness`
```{r}
# Filter the dataset happiness by countries that are in both hemispheres; save that subset in a variable called bothHemispheres and check the first 3 entries
bothHemispheres <- filter(happiness, happiness$Hemisphere == "both")

head(bothHemispheres, 3)
```

```{r}
# Create a new subset of the dataset happiness that includes the countries in which litres of beer consumption per capita are higher than 150; save that in a variable called beer150 and check the first 3 entries
beer150 <- filter(happiness, happiness$Beer_PerCapita > 150)

head(beer150, 3)
```

```{r}
# Create a subset dataset of happiness that only includes from the second to the seventh column, and contains only the first 35 observations; save that subset in a variable called subset1
subset1 <- happiness[1:35, 2:7]

dim(subset1)
```

```{r}
# Create a subset dataset of happiness that includes only the third and sixth column, and contains only the rows with a GDP_PerCapita higher or equal to 40; save it in a variable called subset2and check the first 3 entries

# select the third and sixth column
subset2 <- happiness[, c(3,6)]

# select rows with GDP_PerCapita >= 40
subset2 <- subset(subset2, subset2$GDP_PerCapita >= 40)

head(subset2, 3)
```

`gmd_melt`
```{r}
# Create a subset which only contains NaN values; save that subset as gmd_NaN
gmd_NaN <- gmd_melt[!complete.cases(gmd_melt), ]

# count the number of rows having NaN value
count(gmd_NaN)
```

