---
title: "association_rules_practice"
output: html_notebook
---

# The Brief
By using more than 3 million shopping baskets data provided by **Instacart**, we are going to find the products that re commonly bought together.

# Load Packages
```{r warning=FALSE}
library(arules)
library(arulesViz)
library(dplyr)
```

# Import Data
```{r}
# import the first 20000 rows of orders
orders <- read.csv("order_products__prior.csv", nrows = 20000)

# import all products rows
products <- read.csv("products.csv")
```

```{r}
# check the imported data
head(orders)

head(products)
```

```{r}
# have a glimpse on the data
glimpse(orders)

glimpse(products)
```

The two datasets both have a variable called product_id. So we can join the two datasets on this variable to combine the datasets.

```{r}
# join the two dataset and group them by order id, create a new variable called basket including the products of each order
order_basket <- orders %>%
  inner_join(products, by="product_id") %>%
  group_by(order_id) %>%
  summarise(basket = as.vector(list(product_name)))
```

```{r}
head(order_basket)
```

```{r}
order_basket[1,]$basket
```

```{r}
df <- as(order_basket$basket, "transactions")
head(df)
```

```{r}
dim(df)
```


```{r}
summary(df)
```

```{r}
inspect(df[1:3])
```

```{r}
itemFrequencyPlot(df, topN = 10, type = "absolute", main = "Top 10 Most Frequent Items")
```

### Create Association Rules: Most Common Set of Items Bought
```{r}
# create first association rules
rules <- apriori(df, parameter = list(target="frequent itemsets", support=0.008, minlen=2))

# sort rules by support
rules_support <- sort(rules, by="support", decreasing = TRUE)

# check the first few rules
inspect(rules_support[1:5])
```

```{r}
# use a different support level - 0.01
rules_2 <- apriori(df, parameter = list(target="frequent itemsets", support=0.01, minlen=2))

# sort rules by support
rules_support_2 <- sort(rules_2, by="support", decreasing = TRUE)

# check the most bought itemset
inspect(rules_support_2[1])
```

### Visualize The Most Commom Set of Items Bought
```{r}
rules_df <- DATAFRAME(rules_support)

head(rules_df)
```

```{r}
library(ggplot2)

top_10 <- arrange(rules_df, support)
top_10$items <- factor(top_10$items, levels = top_10$items)
ggplot(top_10, aes(x = items, y = support)) + geom_col() +
  coord_flip()
```

### Create Rules: First set of Association Rules
```{r}
# Create a set of association rules
as_rules <- apriori(df, parameter = list(supp = 0.002, conf = 0.6, minlen=2, maxlen=3))

# Print out the summary of rules
summary(as_rules)
```

```{r}
# Inspect all the 12 rules
inspect(as_rules)
```

```{r}
df_as_rules <- DATAFRAME(as_rules)
head(df_as_rules)
```

```{r}
df_as_rules$rule <- paste(df_as_rules$LHS, " => ", df_as_rules$RHS)
head(df_as_rules)
```


```{r}
# Create a scatterplot with support and confidence on the x and y axis (respectively), and lift as the shading of the dots
plot(as_rules,method = 'scatterplot', measure = c("support","confidence"), shading = "lift")
```

```{r}
# Create a graph plot for the first 10 rules
plot(as_rules, method = "graph", control = list(cex=0.5))
```

```{r}
# Sort the rules by confidence, most confident rules on top and print out the first 10
inspect(sort(as_rules, by = "confidence")[1:10])
```

```{r}
# Inspect the (first few) rules sorted by lift
inspect(sort(as_rules, by = "lift")[1:5])
```

### Study Limes
```{r}
#  Create a subset of the rules called lime_rules and sort it by decreasing lift.
rules = apriori(data=df, parameter=list(support=0.003,confidence=0.2, minlen=2, maxlen=3))

lime_rules = sort(
  subset(rules, subset=rhs %in% 'Limes'),
  by = 'lift',decreasing = T
)
inspect(lime_rules)
```

```{r}
lime_rules_2 <- apriori(df, parameter = list(support=0.003,confidence=0.2, minlen=2, maxlen=3), appearance = list(default = "lhs", rhs = "Limes"))

inspect(lime_rules_2)
```

```{r}
plot(lime_rules_2, method="graph",control = list(cex=0.5))
```

